---
title: "Multiple Regression"
author: "Math 271"
date: "Spring 2022"
output: 
  html_document:
    css: lab.css
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(plotly)

predeqn <- function(model, fmt="%+0.3f"){
  cv <- coef(model)
  cn <- names(cv) %>% str_replace("\\(Intercept\\)", "1")
  cv <- cv %>% sprintf(fmt=fmt)
  paste(cv, "\\cdot \\text{", cn, "}", collapse = " ")
}
```



## One Numerical and One Categorical Predictor

Consider again the `openintro::evals` data set. We will attempt to model the `score` using both `age` and `gender`.

```{r score1}
(score1 <- lm(score ~ age + gender, openintro::evals))
```

From our work last week, we should be able to recognize that the prediction equation for this model will be
\[\widehat{\text{score}} = `r predeqn(score1)`\]

The `gendermale` is a dummy variable which is `1` for males, and `0` for females. 

```{r score1_plot}
evalsplot <- ggplot(openintro::evals) + aes(x=age, y=score, color=gender) + 
  geom_jitter(width=0.5, height=0, alpha=0.5)
ggplotly(evalsplot + geom_line(aes(y=predict(score1))))
```
The ultimate effect of this prediction equation is to produce one regression line for each gender. The predictions for the male group have the coefficient for `gendermale` added, which causes the intercept to change. This is a vertical shift of the regression line.

### Adding an interaction

```{r score2}
(score2 <- lm(score ~ age * gender, openintro::evals))
```
When adding an interaction term (the `*` means add both terms and their interaction), the prediction equation becomes
\[\widehat{\text{score}} = `r predeqn(score2)`\]

Recall that an interaction variable is nothing more than a new variable formed by the product of two others. Since `gendermale` is an indicator variable, the interaction `age:gendermale` has all zeros for females, while males have simply their `age`. By considering the two group equations separately and simplifying, we see that the net effect of this is that now the prediction lines can have different slopes for the two groups.

```{r score2_plot}
ggplotly(evalsplot + geom_line(aes(y=predict(score2))))
```

## Two Numerical Predictors


```{r score3}
(score3 <- lm(score ~ bty_avg + age, openintro::evals))
```

```{r score3_plot}
seq_grid <- function(x, n=10){seq(min(x), max(x), length.out=n)}
newx <- openintro::evals %>% summarize(across(bty_avg|age, seq_grid))
predmat <- newx %>% do.call(expand_grid, .) %>% predict(score3, .) %>% matrix(nrow=nrow(newx))

plotly::plot_ly(openintro::evals, x=~bty_avg, y=~age, z=~score) %>% 
  add_markers(size=1) %>% 
  add_surface(x=~bty_avg, y=~age, z=predmat, data=newx)
```

```{r score4}
(score4 <- lm(score ~ bty_avg * age, openintro::evals))
```

```{r score4_plot}
predmat4 <- newx %>% do.call(expand_grid, .) %>% predict(score4, .) %>% matrix(nrow=nrow(newx))

plotly::plot_ly(openintro::evals, x=~bty_avg, y=~age, z=~score) %>% 
  add_markers(size=1) %>% 
  add_surface(x=~bty_avg, y=~age, z=predmat4, data=newx)
```

