---
title: "Multiple Regression"
author: "Math 271"
date: "Spring 2022"
output: 
  html_document:
    css: lab.css
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(plotly)

predeqn <- function(model, fmt="%+0.3f"){
  cv <- coef(model)
  cn <- names(cv) %>% str_replace("\\(Intercept\\)", "1")
  cv <- cv %>% sprintf(fmt=fmt)
  paste(cv, "\\cdot \\text{", cn, "}", collapse = " ")
}
```



## One Numerical and One Categorical Predictor

Consider again the `openintro::evals` data set. We will attempt to model the `score` using both `age` and `gender`.

```{r score1}
(score1 <- lm(score ~ age + gender, openintro::evals))
```

From our work last week, we should be able to recognize that the prediction equation for this model will be
\[\widehat{\text{score}} = `r predeqn(score1)`\]

The `gendermale` is a dummy variable which is `1` for males, and `0` for females. 

```{r score1_plot}
evalsplot <- ggplot(openintro::evals) + aes(x=age, y=score, color=gender) + 
  geom_jitter(width=0.5, height=0, alpha=0.5)
ggplotly(evalsplot + geom_line(aes(y=predict(score1))))
```
The ultimate effect of this prediction equation is to produce one regression line for each gender. The predictions for the male group have the coefficient for `gendermale` added, which causes the intercept to change. This is a vertical shift of the regression line.

### Adding an interaction

```{r score2}
(score2 <- lm(score ~ age * gender, openintro::evals))
```
When adding an interaction term (the `*` means add both terms and their interaction), the prediction equation becomes
\[\widehat{\text{score}} = `r predeqn(score2)`\]

Recall that an interaction variable is nothing more than a new variable formed by the product of two others. Since `gendermale` is an indicator variable, the interaction `age:gendermale` has all zeros for females, while males have simply their `age`. By considering the two group equations separately and simplifying, we see that the net effect of this is that now the prediction lines can have different slopes for the two groups.

\[\widehat{\text{score}} = 
\begin{cases}
4.88 - 0.018 \cdot \text{age}&\text{ if gender is female}\\
(4.88 - 0.45) + (-0.018 + 0.014)\cdot\text{age}&\text{ if gender is male}
\end{cases}\]

```{r score2_plot}
ggplotly(evalsplot + geom_line(aes(y=predict(score2))))
```

1. Use the code `vole_trapping <- readr::read_csv("http://bit.ly/2IgDF0E")` to import a new data set. Make an initial scatterplot of the `vole_trapping` data, using `veg` as the x-axis, `voles` as the y-axis, and `soil` as the color.

2. Fit a main-effects only model to the vole data and store it as variable `vole_main`. Generate predictions and add the prediction lines to the scatterplot. The lines should be parallel.

3. Fit a model with interaction between the predictors and store it as `vole_inter`. Generate predictions and add the prediction lines to a new scatterplot.

4. Using LaTeX, typeset the prediction equations for both of the above models.

(Data from ModernDive [Lab 06](https://moderndive.github.io/moderndive_labs/static/PS/PS06_multiple_reg.html) on multiple regression.)

## Two Numerical Predictors

The most "natural" way to visualize a regression model using two numeric predictor variables is as a _regression surface_. The predictor variables are placed on the horizontal axes, and the vertical axis holds the value of the prediction. For a main effects only model, the surface will be a flat plane.

```{r score3}
(score3 <- lm(score ~ bty_avg + age, openintro::evals))
```

```{r score3_plot}
seq_grid <- function(x, n=10){seq(min(x), max(x), length.out=n)}
newx <- openintro::evals %>% summarize(across(bty_avg|age, seq_grid))
predmat <- newx %>% do.call(expand_grid, .) %>% predict(score3, .) %>% matrix(nrow=nrow(newx))

plotly::plot_ly(openintro::evals, x=~bty_avg, y=~age, z=~score) %>% 
  add_markers(size=1) %>% 
  add_surface(x=~bty_avg, y=~age, z=predmat, data=newx)
```

Models with interactions have surfaces that can bend into paraboloids or hyperboloids.

```{r score4}
(score4 <- lm(score ~ bty_avg * age, openintro::evals))
```

```{r score4_plot}
predmat4 <- newx %>% do.call(expand_grid, .) %>% predict(score4, .) %>% matrix(nrow=nrow(newx))

plotly::plot_ly(openintro::evals, x=~bty_avg, y=~age, z=~score) %>% 
  add_markers(size=1) %>% 
  add_surface(x=~bty_avg, y=~age, z=predmat4, data=newx)
```

5. Consider the `openintro::mtl` data set. The variable of interest is `total` which measures the thickness of the medial temoral lobe (MTL), along with other information about an individual. In particular, we shall focus on the `age` and `sitting` variables as predictors. Fit a main-effects only model and save it as `mlt_main`. Typeset the resulting prediction equation using LaTeX.

6. (Advanced) Use plotly to create a 3D plot of the regression surface.

7. Fit a model including an interaction term. Typeset and explain the resulting prediction equation. In particular, explain how the interaction term contributes to the prediction.

8. (Advanced) Use plotly to create a 3D plot of the regression surface.

9. Consider the model formula `total ~ age + sitting + sex`. What columns are in the model matrix? Describe the prediction equation for this model.

10. (Advanced) Try out the models `total ~ age * sitting + sex` and `total~age * sitting * sex` and describe the model matrices and prediction equations for these models.


## Plotly

The [`plotly` library](https://plotly.com/r/) is a library which produces ineractive graphics for html output formats. The `ggplotly` command in particular will attempt to convert a ggplot object into a plotly graphic, and is a simple way to start using the library. It can also be used directly, with the `plot_ly` command.